Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming

### 2017-05-11 docs/paperin.tex
move design var table; tweak abstract, intro, and conclusion

\begin{abstract}
    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
    These aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, wind speeds, operational capability and coverage footprint.
    In this paper, the feasible limits of solar-elecrtic and gas powered, long-endurance aircraft are analyzed for various requirements and design variables to be used for initial vehicle sizing and design. 
    Additionally, this paper demonstrates the capability of geometric programming, a form of convex optimization, as a tool to evaluate the design trade offs between architectures and requirements.
    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
\end{abstract}

@@

This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.  Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.

@@

Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
Varying mission requirements reveals how the best architecture changes depending on the requirements.

@@

The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.

\newpage
\begin{longtable}{llll}
\caption{Fixed Optimization Parameters} \\
\toprule
\toprule
\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
\midrule
$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
\bottomrule
\label{t:gassolarparams}
 \end{longtable}

@@

When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.  
Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.

@@

\input{sens.generated.tex}

\newpage

\input{svals.generated.tex}

\input{gassens.generated.tex}

\input{gvals.generated.tex}

@@

Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.  
One high level discovery is the importance of wind speed.  
If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.  Using higher energy density batteries can result in significant weight and performance savings.  
The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.

>>>>

\begin{abstract}
    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design. 
    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
\end{abstract}

@@

This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.  Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
The optimization can be solved in a fraction of a second and is used to the feasible limits of the solar-electric and gas powered, long-endurance aircraft.

@@

Using this geometric programming methodology, trade offs between different aircraft configurations, power sources, and requirements can be calculated in seconds.  
The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
Varying mission requirements reveals how the best architecture changes depending on the requirements.

@@

The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs} and the assumptions and key design parameters described herein. 
The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
Key design variables for both architectures are listed in Tables~\ref{t:svals} and~\ref{t:gvals} for various mission requirements. 
The rest of this section explores how these result may vary for different requirements, input parameter values, and physics modeling assumptions.

\newpage
\input{svals.generated.tex}
\input{gvals.generated.tex}
% \begin{longtable}{llll}
% \caption{Fixed Optimization Parameters} \\
% \toprule
% \toprule
% \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
% \midrule
% $\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
% $\tau$                       & 0.115 & $\tau$                       & 0.115 \\
% $\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
% $\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
% $N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
% $N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
% $\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
% Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
% $\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
% $\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
%                              &     & $\eta_{\text{discharge}}$       & 0.98 \\
% \bottomrule
% \label{t:gassolarparams}
%  \end{longtable}

@@

When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.  
Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.

@@

\input{sens.generated.tex}
\input{gassens.generated.tex}

@@

Using geometric programming the feasibility limits and design space of solar-electric and gas powered aircraft is analyzed is second to a level of detail and speed not previously achieved in initial aircraft sizing and design. 
Based on the assumptions included herein, the gas powered aircraft is more capable of meeting high latitude and station keeping requirements. 
The solar-electric powered aircraft sizing does not converge at latitude requirements above 31 degrees latitude. 
However, solar-electric powered aircraft can meet high endurance and altitude requirements.  
The effect of key design parameters on the size of each architecture was also achieved.  
Using higher energy density batteries can result in significant weight and performance savings for the solar-electric powered aircraft.  
The rapid solve time of geometric programming allows detailed sizing and design analysis of long-endurance aircraft and the capability of solar-electric and gas powered aircraft to meet various requirements.
